---
title: "Using riverbed with multiple channels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multiple_channels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, message=FALSE, warning=FALSE
)
```

```{r setup}
library(riverbed)
library(dplyr)
library(ggplot2)
```

# Load example datasets


Use two transects s1 and s2 **already included as example datasets** in the package. They correspond to the same place, at two different times.

```{r data_load}
data(s1)
data(s2)
```

You can  compare those two transects and assess the overall change between them through a simple call to `area_between()`. 

```{r simple_channel_calc}
result_area=area_between(s1,s2)
result_area
plot_area(result_area)
```

# Identify channels with get_channels()

On the other hand, these transects might correspond to a multi-channel riverbed, in which case you might want to assess the changes of each channel individually. It is then necessary to first **identify channels**.


```{r identify_channels1, fig.width=6,fig.height=3}
channels1=get_channels(s1, hmin=1, hmax=5.5)
channels1
plot_channels(channels1,sr=s1)
```

- i_a and i_b: indices of observations delimiting the channel,
- l_a and l_b: longitudinal coordinates of points delimiting the channel
- z_a and z_b: height of points delimiting the channel
- z_min and z_max: minimum and maximum height of the channel-delimited profile

The type of channel returned can be either

- "bankfull" (the default, as above)
- "levee-to-levee" (see example below), which is delimited on both sides by local maxima

Here is what you would get with type="levee-to-levee":

```{r identify_channels1_levee_to_levee, fig.width=6,fig.height=3}
channels1=get_channels(s1, hmin=1, hmax=5.5, type="levee-to-levee")
channels1
plot_channels(channels1,sr=s1)
```

# Effect of parameters hmin and hmax

**Parameters hmin and hmax**

**Local minima** must be at $z<z_{min}+h_{max}$ to define a channel's bottom.

**Local maxima** must be at $z>z_{localmin}+h_{min}$ to define a channel's banks.

Hence $hmin$ corresponds to the minimum water depth in a channel, and $hmax$ corresponds to the maximum height of a channel's bottom.

See the effect of these parameters:

```{r test2, fig.width=6,fig.height=3}
channels1=get_channels(s1, hmin=0.5, hmax=4)
plot_channels(channels1,sr=s1)
```

```{r test3, fig.width=6,fig.height=3}
channels1=get_channels(s1, hmin=3, hmax=5)
plot_channels(channels1,sr=s1)
```
# Regroup channels through dates

Here, with the same values $h_{min}$ and $h_{max}$, we get **3 distinct channels at both dates**:

```{r two_channels, fig.width=6, fig.height=3}
channels1=get_channels(s1,hmin=1,hmax=5.5)
plot_channels(channels1,sr=s1)
channels2=get_channels(s2,hmin=1,hmax=5.5)
plot_channels(channels2,sr=s2)
```

... But these correspond to different $l$ coordinates.

So, we have to somehow **"regroup" these coordinates** to be able to compare the transects with **channel-wise area calculations**.

# Example of channel-wise area calculation through two dates


## get_channels()

Let's apply the `get_channels()` function to all transects (the column identifying those being called "id_transect").

```{r gather_transects}
transects=bind_rows(bind_cols(id_transect="s1",s1),
                    bind_cols(id_transect="s2",s2))%>% 
  group_by(id_transect) %>% 
  tidyr::nest()
```

The table `transects` gathers the data of both transects s1 and s2. Let's apply `get_channels()` iteratively to each transect:

```{r get_all_channels}
channels =transects %>%  
  mutate(data=purrr::map(data,get_channels, hmin=1, hmax=4.5)) %>% 
  tidyr::unnest(cols=data) 
channels
```
## regroup_channels()

Both transects correspond to **3 channels** with identifier `id_c` (1,2,3).

We regroup these channels with function `regroup_channels()`, which returns for each channel longitudinal limits `l_a` and `l_b`. These limits include the total width of both channels.

```{r regroup_channels, fig.width=6, fig.height=3}
regrouped_channels=regroup_channels(channels,by_id=id_c)
regrouped_channels
plot_channels(channels1, s1)+
  geom_vline(data=regrouped_channels, aes(xintercept=l_a), col="grey")+
  geom_vline(data=regrouped_channels, aes(xintercept=l_b), col="grey")
plot_channels(channels2, s2)+
  geom_vline(data=regrouped_channels, aes(xintercept=l_a), col="grey")+
  geom_vline(data=regrouped_channels, aes(xintercept=l_b), col="grey")
```
## area_between()

```{r calc_areas_result}
areas_result=transects %>%
  mutate(subdata=purrr::map(data,cut_series,
                            channels=regrouped_channels)) %>% 
  tidyr::unnest(cols=subdata) %>% 
  tidyr::pivot_wider(id_cols=id_c, names_from=id_transect,values_from=sr) %>%
  mutate(area_complete=purrr::map2(s1,s2,area_between))
areas_result
```

## format result

```{r format_areas_result}
areas_formatted_result= areas_result%>%
  mutate(area=purrr::map_dbl(area_complete,~.x$area)) %>% 
  mutate(area_lower=purrr::map_dbl(area_complete,
                                   ~.x$area_by_type %>%
                                     filter(type=="lower") %>% pull(area))) %>% 
  mutate(area_upper=purrr::map_dbl(area_complete,
                                   ~.x$area_by_type %>% 
                                     filter(type=="upper") %>% pull(area))) %>% 
  mutate(plot=purrr::map(area_complete,plot_area)) %>% 
  select(id_c,area,area_lower,area_upper,plot)
areas_formatted_result
```

```{r plot_areas, fig.width=3, fig.height=3}
areas_formatted_result$plot
```



