---
title: "A quick glimpse of riverbed"
author: "L. Vaudor"
date: "10 juillet 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(riverbed)
s1 <- tibble(l=c(1.1,3.4,5.0,5.3,6.2,4.3,2.5,6.7,6.9,9.2),
             z=c(1.4,3.1,5.5,4.8,1.3,2.5,4.3,2.2,3.1,2.6))
s2 <- tibble(l=c(0.5,2.5,4.0,6.0,4.4,5.8,9.5,7.2),
             z=c(2.8,1.5,2.2,4.3,5.5,2.2,1.5,5.3))
```

We consider two longitudinal series $s_1=(l_1,z_1)$, and $s_2=(l_2,z_2)$:

```{r, echo=FALSE, fig.width=5, fig.height=3}
result_area <- area_between(s1,s2)

ggplot(filter(result_area$data, !is.na(series)),
       aes(x=l_obs,y=z_obs,color=series)) +
  geom_line()+
  geom_point()
```

To calculate the area between the curves we need to interpolate the data and calculate intersects between the two curves:

```{r, echo=FALSE, fig.width=5, fig.height=3}
ggplot(result_area$data,
       aes(x=l_obs,y=z_obs,color=series)) +
  geom_line()+
  geom_point()+
  geom_point(data=filter(result_area$data,
                         p=="interpolated",
                         series=="s1"),
             aes(x=l,y=z1),shape="|",size=3)+
  geom_point(data=filter(result_area$data,
                         p=="interpolated",
                         series=="s2"),
             aes(x=l,y=z2),shape="|",size=3)+
  geom_point(data=filter(result_area$data,
                         p=="intersect"),
             aes(x=l,y=z1), col="dark blue",shape=0)
```

This, as well as the calculation of corresponding trapezia, is realised by the function `area_between()`.

```{r}
result_area <- area_between(s1,s2)
result_area
```

The output of the function corresponds to 

- **total area** between curves (i.e. positive area - negative area)
- area between curves according to whether first curve is above the second curve ("**upper**") or below ("**lower**")
- data with all **coordinates** of interpolated points and intersects, used to calculate the area of trapezia (and which will be used by function `plot_area()`)

Note that it is also possible to calculate the area between one longitudinal series $s_1=(l_1,z_1)$ and a constant horizontal value $h$.

```{r}
result_area_h <- area_between(s1,h=2)
```

## Plot of area

```{r, fig.width=5, fig.height=3}
plot_area(result_area)
```

If the distinction between upper and lower area is irrelevant then it is possible **not to distinguish these two types** on the plot:

```{r, fig.width=5, fig.height=3}
plot_area(result_area,distinguish_type=FALSE)
```

The plotting function `plot_area()` can also be used with just **one series** and **constant horizontal level** $h$, as calculated earlier:

```{r plot_area_h, fig.width=5, fig.height=3}
plot_area(result_area_h)
```

# Error in the calculation of area

Now suppose that the series $s_1$ and $s_2$ are measured with a certain level of uncertainty, due to 

- imprecision in the measure of height $z$ ($\sigma_z$) and/or 
- imprecision in the measure of longitudinal coordinate $l$ ($\sigma_l$). 

These imprecisions might be different for the two series (change in sampling protocol, improvement in sampling gear between two dates, etc.).


The **uncertainty in measures** results in a certain amount of **uncertainty in the estimate of area** between the two curves. This is provided by the function `area_uncertainty()`.

Here, with errors in the measures of height (0.1 and 0.3) higher than in the measures of longitudinal coordinates (0.05 and 0.2), and errors more important for series $s_2$ (0.3 and 0.2) than for $s_2$ (0.1 and 0.05).

```{r result_area_with_errors}
result_area <- area_between(s1,s2,
                            sigma_z=c(0.1,0.3),
                            sigma_l=c(0.05,0.2))
```

The `plot_area()` function can provide a visual hint of the uncertainties in $l$ and $z$ measures with horizontal and vertical error bars:

```{r show_uncertainty, fig.width=5,fig.height=3}
plot_area(result_area, show_uncertainty=TRUE)
```

```{r estimates_and_confidence_interval, echo=FALSE}
A_est <- result_area$area
sigma_area <- round(result_area$sigma_area,2)
A_liminf <- round(A_est-1.96*A_est,2)
A_limsup <- round(A_est+1.96*A_est,2)
```

Here, for instance, 

- The estimate of area is `r round(A_est,2)`.
- The uncertainty in measures results in an error in the estimate of area of $\sigma_{area}$=`r sigma_area`. Hence we have a 95% confidence interval for the estimate of area of $[A-1.96\cdot\sigma_{area}, A+1.96\cdot\sigma_{area}]$=[`r A_liminf`,`r A_limsup`].

We can check the calculation of error through a simple simulation with 1000 series $s_{1tmp}$ and $s_{2tmp}$ varying around $s_1$ and $s_2$ respectively, with variations corresponding to estimation errors $\sigma_z$ and $\sigma_l$:

```{r, cache=TRUE}
set.seed(33)
sigma_z=c(0.1,0.3)
sigma_l=c(0.05,0.2)
f=function(i){
  s1_tmp <- tibble(l=s1$l+rnorm(nrow(s1),0,sigma_l[1]),
                   z=s1$z+rnorm(nrow(s1),0,sigma_z[1]))
  s2_tmp <- tibble(l=s2$l+rnorm(nrow(s2),0,sigma_l[2]),
                   z=s2$z+rnorm(nrow(s2),0,sigma_z[2]))
  return(area_between(s1_tmp,s2_tmp)$area)
}
area_vals=purrr::map_dbl(1:1000,f)
sd(area_vals)

res=area_between(s1,s2,
                 sigma_z=sigma_z, sigma_l=sigma_l)
res$sigma_area
```

