---
title: "riverbed"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riverbed}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, message=FALSE, warning=FALSE
)
```

```{r setup}
library(riverbed)
library(dplyr)
library(ggplot2)
```

# Example datasets

```{r data_load}
data(s1)
data(s2)
```

We consider two longitudinal series $s_1=(l_1,z_1)$, and $s_2=(l_2,z_2)$:

```{r plot_series, echo=FALSE, fig.width=5, fig.height=3}
result_area <- area_between(s1,s2)

ggplot(filter(result_area$data, !is.na(series)),
       aes(x=l_obs,y=z_obs,color=series)) +
  geom_line()+
  geom_point()
```

# Principle of the area calculation

To calculate the area between the curves we need to interpolate the data and calculate intersects between the two curves:

```{r plot_points_types, echo=FALSE, fig.width=5, fig.height=3}
ggplot(result_area$data,
       aes(x=l_obs,y=z_obs,color=series)) +
  geom_line()+
  geom_point()+
  geom_point(data=filter(result_area$data,
                         p=="interpolated",
                         series=="s1"),
             aes(x=l,y=z1),shape="|",size=3)+
  geom_point(data=filter(result_area$data,
                         p=="interpolated",
                         series=="s2"),
             aes(x=l,y=z2),shape="|",size=3)+
  geom_point(data=filter(result_area$data,
                         p=="intersect"),
             aes(x=l,y=z1), col="dark blue",shape=0)
```

# Use of area_between()

This, as well as the calculation of corresponding trapezia, is realised by the function `area_between()`.

```{r result_area}
result_area <- area_between(s1,s2)
result_area
```

The output of the function corresponds to 

- **total area** between curves (i.e. positive area - negative area)
- area between curves according to whether first curve is above the second curve ("**upper**") or below ("**lower**")
- data with all **coordinates** of interpolated points and intersects, used to calculate the area of trapezia (and which will be used by function `plot_area()`)

Note that it is also possible to calculate the area between one longitudinal series $s_1=(l_1,z_1)$ and a constant horizontal value $h$.

```{r result_area_h}
result_area_h <- area_between(s1,h=225)
```

# Plot of area

```{r plot_area, fig.width=5, fig.height=3}
plot_area(result_area)
```

If the distinction between upper and lower area is irrelevant then it is possible **not to distinguish these two types** on the plot:

```{r plot_area_not_distinguished, fig.width=5, fig.height=3}
plot_area(result_area,distinguish_type=FALSE)
```

The plotting function `plot_area()` can also be used with just **one series** and **constant horizontal level** $h$, as calculated earlier:

```{r plot_area_h, fig.width=5, fig.height=3}
plot_area(result_area_h)
```
